---
title: "R Notebook"
output:
  html_document:
    df_print: paged
---

```{r}
library(dplyr)
library(gapminder)
data("gapminder")
attach(gapminder)
getwd()
gapminder <- read.csv("gapminder_data.csv", header=TRUE)
head(gapminder)
```

I've loaded the gapminder data and now I am goint to manipulate it with dplyr.

```{r}
gap <- gapminder
rm(gapminder)
mean(gap[gap$continent == "Africa", "gdpPercap"])
mean.gdpAfrica <- mean(gap$gdpPercap[continent=="Africa"])
mean.gdpAfrica
mean.gdpAsia <- mean(gap$gdpPercap[continent=="Asia"])
mean.gdpAsia
mean.gdpAmericas <- mean(gap$gdpPercap[continent=="Americas"])
mean.gdpAmericas
```
### working with data the dplyr way

select() - subset variables    (columns)
filter() - subset observations   (rows)
group_by() - counts and groups data
summarize() - useful information about the data
mutate() - add a new column based on the data

```{r}
ncol(gap)

names(gap)

subset_gap <- gap %>% select(year, country, gdpPercap)

subset_gap
dim(subset_gap)
```

filter() : 행을 바꿈
filter먼저하고 select해야함
%>% means permutation(치환)

```{r}
year_country_gdp_euro <- gap %>%
  filter(continent == "Europe") %>%
  select(year,country,gdpPercap)

year_country_gdp_euro

```

### Quiz 1

```{r}
str(gap)
library(gapminder)
gapminder <- read.csv("gapminder_data.csv", header=TRUE)
answer <- gapminder %>%
  filter(continent == "Africa") %>%
  filter(year >= 2007) %>%
  select(year,country,lifeExp)
nrow(answer)

```

### Possible answer 1

```{r}
answer <- gapminder %>%
  filter(continent == "Africa", year == 2007) %>%
  select(year,country,lifeExp)

```

group_by() function subset your data by chunking it based on a column

```{r}
head(gap)

str(gap)

gap_by_continent <- gap %>% group_by(continent)
head(gap_by_continent)
str(gap_by_continent)

```

group_by() turns a single data frame into a set of data frames grouped based on some variable.
```{r}
gdp_by_continet <- gap_by_continent %>%
  summarize(mean_gdpPercap = mean(gdpPercap))
gdp_by_continet

gdp <- gap %>%
  summarize(mean_gdpPercap = mean(gdpPercap))

gdp


```

### #Quiz 2

```{r}
group_by_year <- gap %>% 
  filter(continent == "Africa") %>% 
  group_by(year) %>% 
  summarize(mean_lifeExp = mean(lifeExp))
group_by_year

```

Calculate the average life expectancy per country. Which has the logest average life expectancy? and Which has the shortest averge life expectancy?

```{r}
life <- gap %>%
  group_by(country) %>%
  summarize(mean_lifeExp = mean(lifeExp))
life
life %>% filter(mean_lifeExp == min(mean_lifeExp))
life %>% filter(mean_lifeExp == max(mean_lifeExp))

```

mutate() is a function in dplyr that lets you add a column to a data frame based the data.

```{r}
billgdp <- gap %>%
  filter(year == 2007) %>%
  mutate(billion_gdp = gdpPercap * pop / 10^9) %>%
  select(continent, country, billion_gdp)

colnames(billgdp)
dim(billgdp)

```

```{r}
gdp_by_cont <- gap %>%
  group_by(continent, year) %>%
  summarize(mean_gdp = mean(gdpPercap),
            sd_gdp = sd(gdpPercap),
            mean_pop = mean(pop),
            sample_size = n(), #function n() counts sample size of the var grouped by
            se_gdp = sd_gdp / sqrt(sample_size))
head(gdp_by_cont)
str(gdp_by_cont)
gdp_by_cont <- gdp_by_cont %>% data.frame()
```

### #Combining dplyr with ggplot2
```{r}
library(ggplot2)
gap %>% filter(continent == "Asia") %>%
  ggplot(aes(x = gdpPercap, y = lifeExp)) +
  geom_point()
```

```{r}

gap_wide <- read.csv("gapminder_wide.csv", stringsAsFactors = FALSE)
dim(gap_wide)
colnames(gap_wide)

```

This data is in a wide format. we want to change it to a long format.
```{r}
head(gap_wide)

## To make the data long, we will use the function gather()
library(tidyr)
gap_long <- gap_wide %>%
  gather(obstype_year, obs_values, -continent, -country)
dim(gap_long)
head(gap_long)
```

## 엑셀에서 텍스트 나누기 느낌
## sep 함수
```{r}
gap_long <- gap_long %>% separate(obstype_year, into = c("obs_type", "year"), sep="_")
head(gap_long)
```

## 위에서 분리한 year 열을정수로 정의하기

```{r}
gap_long$year <- as.integer(gap_long$year)
```

### Quiz 4

```{r}
head(gap_long)
gap_long %>% filter(obs_type == "lifeExp") %>%
  group_by(continent) %>%
  summarize(mean_lifeExp = mean(obs_values))

str(gap_long)
```
### Changing to human readable format

```{r}
gap_normal <- gap_long %>% spread(obs_type, obs_values)
str(gap_normal)
colnames(gap_normal)

str(gap)
colnames(gap)
# 원래 데이터(gap) 열 순서로 바꾸기
gap_normal <- gap_normal[,names(gap)]
names(gap)
names(gap_normal)

# 두 데이터 비교해보기
all.equal(gap, gap_normal)
head(gap)
head(gap_normal)

gap_normal <- gap_normal %>% arrange(country, continent, year)
head(gap_normal)
all.equal(gap, gap_normal)

str(gap)
str(gap_normal)
```



